---
title: "cm009 Notes and Exercises: Table Joins"
date: '2017-10-03'
output: github_document
---

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gapminder))
```

## notes from class
```{r}
library("nycflights13")
# Drop unimportant variables so it's easier to understand the join results.
flights2 <- flights %>% select(year:day, hour, origin, dest, tailnum, carrier)
flights2 %>% 
  left_join(airlines)
#> Joining, by = "carrier"
#> # A tibble: 336,776 x 9
#>    year month   day  hour origin dest  tailnum carrier name               
#>   <int> <int> <int> <dbl> <chr>  <chr> <chr>   <chr>   <chr>              
#> 1  2013     1     1  5.00 EWR    IAH   N14228  UA      "United Air Lines …
#> 2  2013     1     1  5.00 LGA    IAH   N24211  UA      "United Air Lines …
#> 3  2013     1     1  5.00 JFK    MIA   N619AA  AA      "American Airlines…
#> 4  2013     1     1  5.00 JFK    BQN   N804JB  B6      "JetBlue Airways"  
#> 5  2013     1     1  6.00 LGA    ATL   N668DN  DL      "Delta Air Lines I…
#> # ... with 3.368e+05 more rows

```

Notes: left-join will match info from the first database with the second one by a matching variable, and will add the rest of the columns that are not in the other one.
It only keeps the original dataframe

inner-join will match by common rows (eliminate the other ones)
outer-join wlll keep rows from both dataframes
check on Jenny Bryan examples [here](http://stat545.com/bit001_dplyr-cheatsheet.html)



After going through the `dplyr` [vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/two-table.html) on "two-table verbs", we'll work on the following exercises.


Consider the following areas of countries, in hectares:

```{r}
(areas <- data.frame(country=c("Canada", "United States", "India", "Vatican City"),
                     area=c(998.5*10^6, 983.4*10^6, 328.7*10^6, 44)) %>% 
     as.tbl)
```


1. To the `gapminder` dataset, add an `area` variable using the `areas` tibble. Be sure to preserve all the rows of the original `gapminder` dataset.

```{r}
gapminder%>% 
  left_join(areas)
```

2. To the `gapminder` dataset, add an `area` variable using the `areas` tibble, but only keeping obervations for which areas are available. 

```{r}
gapminder %>% 
  inner_join(areas)  #only rows
```

3. Use a `_join` function to output the rows in `areas` corresponding to countries that _are not_ found in the `gapminder` dataset. 

```{r}
anti_join(areas, gapminder)
```


4. Use a `_join` function to output the rows in `areas` corresponding to countries that _are_ found in the `gapminder` dataset. 
```{r}
semi_join(areas, gapminder)
```

5. Construct a tibble that joins `gapminder` and `areas`, so that all rows found in each original tibble are also found in the final tibble. 
```{r}
full_join(gapminder, areas)
```

6. Subset the `gapminder` dataset to have countries that are only found in the `areas` data frame. 

```{r}
semi_join(gapminder, areas)
```


7. Subset the `gapminder` dataset to have countries that are _not_ found in the `areas` data frame. 
```{r}
anti_join(gapminder, areas)
```
```{r}
anti_join(gapminder, areas) %>% 
  filter(country == "Canada")
```

