---
title: "cm010 - cm11 Notes"
author: "Santiago D"
date: '2017-10-05'
output: github_document
---

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gapminder))
```

Different types of joins (from previous class): `left_join`,`inner_join`...etc

```{r}
install.packages("devtools") #if you dont have it
library(devtools)
install_github("JoeyBernhardt/singer")
library(singer)
```

Load singer

```{r}
data("songs")
data("locations")
glimpse(locations)
glimpse(songs)
```

### get a dataframe with release and year
get a dataframe with all albums, artists and the year in which album was published
```{r}
songs %>% 
  right_join(locations) %>% 
  select(release, artist_name, year)

# We can just try all
relartyear <- inner_join(locations, songs, by = "title") %>% 
  select(release, artist_name.x, year)
View(relartyear)
```

we can also join with artist name so this variable is not duplicate
```{r}
relartyear <- inner_join(locations, songs, by = c("artist_name","title")) %>% 
  select(release, artist_name, year)
View(relartyear)
```

### exercise
Try to get the number of releases per year

```{r}
relartyear %>% 
  group_by(year) %>% 
  summarise(release.per.year = count(year))

inner_join(songs,locations, by = "title") %>% 
  count(year)

```
note: inner_join is symmetric, so doesn't matter which is the order

### RESHAPING

```{r}
data("singer_locations")
glimpse(singer_locations)
?singer_locations
View(singer_locations)
```

**WIDE** dataframe vs **LONG** dataframe

Lets get year, artist_hotttnesss, artist_familiarity, duration
```{r}
hfd_y <- singer_locations %>% 
  select(year, artist_hotttnesss, artist_familiarity, duration)

hfd_y %>% 
  filter(year > 1900) %>% 
  ggplot(aes(x = year, y = artist_hotttnesss)) +
  geom_point()
```

package `tidyr`, `gather()`: from wide to long

```{r}
hfd_y_long <- hfd_y %>% 
  gather(key = "Measure", value = "Units", artist_hotttnesss:duration)
View(hfd_y)
View(hfd_y_long)
```
now, we graph from this long format
```{r}
plot3 <- hfd_y_long %>%
  filter(year > 1900) %>% 
  ggplot(aes(year, Units)) +
  geom_point() +
  facet_wrap(~ Measure, scales = "free_y")  #find a way to display the scale better
```

### more notes from oct 10

```{r}
hfd_y <- singer_locations %>%
  select(year, duration, artist_familiarity, artist_hotttnesss) %>% 
  gather(key = "Measures", value = "My_value",
         duration:artist_familiarity)
```

### Doomed challenge 1
Bring hfd_y in the wide format

```{r}
hfd_y %>% 
  spread(Measures, My_value)
```

We can't spread this info, because years have different values for the variables, so spread doesn't know what to do with those values


### challenge 2 fix hdf_y to include identifiers

```{r}
hfd_y <- singer_locations %>%
  select(year, song_id, duration, artist_familiarity, artist_hotttnesss) %>% 
  gather(key = "Measures", value = "My_value",
         duration:artist_familiarity)

hfd_y %>% 
  spread(Measures, My_value)
```

Now, we try reshape2

```{r}
install.packages("reshape2")
library(reshape2)
```

reshape has other related functions to do the same tasks

`dcast` is another function that allows you to calculate summarize stats from the groups of values

```{r}
hfd_y_mean <- hfd_y %>% 
      dcast(year ~ Measures, 
        value.var = "My_value",
        fun.aggregate = mean, na.rm = TRUE)
View(hfd_y_mean)
```

### Now, getting back to concepts
R has many types of objects in its environment

class vs type

- functions
- lists
- dataframe
- character
- double
- numeric (class)
- 

### Challenge 3: what is a plot
```{r}
typeof(plot3)
class(plot3)
```

### Challenge 4 what is a pipe %>% 
```{r}
typeof(`%>%`)  # you have to use "comillas"
class(`%>%`)

class(`<-`)
`<-`(y,13)
y

```

