---
title: "cm102"
author: "Santiago David"
date: '2017-10-26'
output: github_document
---

```{r}
library(gapminder)
```

```{r}
## get to know the functions mentioned above
min(gapminder$lifeExp)
## [1] 23.599
max(gapminder$lifeExp)
## [1] 82.603
range(gapminder$lifeExp)
## [1] 23.599 82.603

## some natural solutions
max(gapminder$lifeExp) - min(gapminder$lifeExp)
## [1] 59.004
with(gapminder, max(lifeExp) - min(lifeExp))
## [1] 59.004
range(gapminder$lifeExp)[2] - range(gapminder$lifeExp)[1]
## [1] 59.004
with(gapminder, range(lifeExp)[2] - range(lifeExp)[1])
## [1] 59.004
diff(range(gapminder$lifeExp))
## [1] 59.004
```

Functions
We create a function
```{r}
max_minus_min <- function(x) max(x) - min(x)
max_minus_min(gapminder$lifeExp)
```

```{r}
max_minus_min(1:10)
max_minus_min(runif(1000))
```

```{r}
max_minus_min(gapminder$gdpPercap)
max_minus_min(gapminder$pop)
```

stopifnot is an option to constrain the function to specifc rules
```{r}
mmm <- function(x) {
  stopifnot(is.numeric(x))
  max(x) - min(x)
}

mmm(gapminder)
mmm(gapminder$country)
mmm("eggplants are purple")
```

Wecan also create a customized error message
```{r}
mmm2 <- function(x) {
  if(!is.numeric(x)) {
    stop('I am so sorry, but this function only works for numeric input!\n',
         'You have provided an object of class: ', class(x)[1])
  }
  max(x) - min(x)
}
mmm2(gapminder)
```

### Problem 1
1. create a new function mod_ten that takes a number and returns the remainder after dividing by ten. for example mod_ten(15) returns 5. mod_ten(1001) should return 1


```{r}
mod_ten <- function(x) {
  if(!is.numeric(x)){
    stop('I am so sorry, but this function only works for numeric input!\\n',
          'You have provided an object of class: ', class(x)[1])
  }
  x %% 10
}

mod_ten(15)
mod_ten(1001)
```

Generalize our function to other quantiles
```{r}
mmm <- function(x) {
  stopifnot(is.numeric(x))
  max(x) - min(x)
}
```

quantile function in r
```{r}
quantile(gapminder$lifeExp)
quantile(gapminder$lifeExp, probs = 0.5) #50% 
quantile(gapminder$lifeExp, probs = c(0.25, 0.75)) #25% and 75%
boxplot(gapminder$lifeExp, plot = FALSE)$stats
```

we write a snippet
```{r}
the_probs <- c(0.25, 0.75)
the_quantiles <- quantile(gapminder$lifeExp, probs = the_probs)
max(the_quantiles) - min(the_quantiles)
```

Now a function
```{r}
qdiff1 <- function(x, probs) {
  stopifnot(is.numeric(x))
  the_quantiles <- quantile(x = x, probs = probs)
  max(the_quantiles) - min(the_quantiles)
}
qdiff1(gapminder$lifeExp, probs = c(0.25, 0.75))
IQR(gapminder$lifeExp) 
qdiff1(gapminder$lifeExp, probs = c(0, 1))
mmm(gapminder$lifeExp)
```

#### argument names: freedom and conventions
```{r}
qdiff2 <- function(zeus, hera) {
  stopifnot(is.numeric(zeus))
  the_quantiles <- quantile(x = zeus, probs = hera)
  return(max(the_quantiles) - min(the_quantiles))
}
```

# but maybe better to have meaningful names
```{r}
qdiff3 <- function(my_x, my_probs) {
 stopifnot(is.numeric(my_x))
  the_quantiles <- quantile(x = my_x, probs = my_probs)
  return(max(the_quantiles) - min(the_quantiles))
}
qdiff3(my_x = gapminder$lifeExp, my_probs = 0:1)

```

Return

```{r}
return(max(the_quantiles) - min(the_quantiles))
```

default values

```{r}
qdiff1(gapminder$lifeExp)
```

```{r}
qdiff4 <- function(x, probs = c(0, 1)) {
  stopifnot(is.numeric(x))
  the_quantiles <- quantile(x, probs)
  return(max(the_quantiles) - min(the_quantiles))
}
```

checking how the function works
```{r}
qdiff4(gapminder$lifeExp)
mmm(gapminder$lifeExp)
qdiff4(gapminder$lifeExp, c(0.1, 0.9))
```

### how to return multiple values = creating a list

### Problem part 2

1- create general function that takes two arguments, value and divisor
2- set the default parameter of divisor to be 10

```{r}
mod_x <- function(value, divisor = 10) {
  if(!is.numeric(value)){
    stop('not valid input')
  }
  value %% divisor
}

mod_x(15,6)

```

```{r}
#class solution
mod_x <- function(value, divisor = 10){
  stopifnot(!is.integer(divisor))
  int <- value %/% divisor
  mod <- value %% divisor
  #return(c(int,mod))
  output <- list(quotient=int, remainder=mod)
  return(output)
}
mod_object = mod_x(27)
mod_object$quotient
mod_object$remainder
```

### restore max min
```{r}
qdiff4 <- function(x, probs = c(0, 1)) {
  stopifnot(is.numeric(x))
  the_quantiles <- quantile(x, probs)
  return(max(the_quantiles) - min(the_quantiles))
}
```

### NA's
```{r}
z <- gapminder$lifeExp
z[3] <- NA
quantile(gapminder$lifeExp)
quantile(z)
quantile(z, na.rm = TRUE)
```

we can include an argument for the na's in our function, either inside the function, or as a default

```{r}
qdiff5 <- function(x, probs = c(0, 1)) {
  stopifnot(is.numeric(x))
  the_quantiles <- quantile(x, probs, na.rm = TRUE)
  return(max(the_quantiles) - min(the_quantiles))
}
qdiff6 <- function(x, probs = c(0, 1), na.rm = TRUE) {
  stopifnot(is.numeric(x))
  the_quantiles <- quantile(x, probs, na.rm = na.rm)
  return(max(the_quantiles) - min(the_quantiles))
}
qdiff5(gapminder$lifeExp)
```

`... argument`
```{r}
qdiff7 <- function(x, probs = c(0, 1), na.rm = TRUE, ...) {
  the_quantiles <- quantile(x = x, probs = probs, na.rm = na.rm, ...)
  return(max(the_quantiles) - min(the_quantiles))
}

set.seed(1234)
z <- rnorm(10)
quantile(z, type = 1)
quantile(z, type = 4)
all.equal(quantile(z, type = 1), quantile(z, type = 4))
```

### test_that
```{r}
#install.packages("testthat")
#library(testthat)
test_that('invalid args are detected', {
  expect_error(qdiff7("eggplants are purple"))
  expect_error(qdiff7(iris))
  })
test_that('NA handling works', {
  expect_error(qdiff7(c(1:5, NA), na.rm = FALSE))
  expect_equal(qdiff7(c(1:5, NA)), 4)
})
```


TEsting
```{r}
qdiff_no_NA <- function(x, probs = c(0, 1)) {
  the_quantiles <- quantile(x = x, probs = probs)
  return(max(the_quantiles) - min(the_quantiles))
}
test_that('NA handling works', {
  expect_that(qdiff_no_NA(c(1:5, NA)), equals(4))
})
```

