---
title: "cm104_purrr_mynotes"
author: "Santiago David"
date: '2017-11-02'
output: github_document
---

### Nested information, lists...

nested lists are complicated objects
```{r}
library(tidyverse)
library(purrr)
library(stringr)
library(repurrrsive)
library(listviewer)
library(jsonlite)
library(tibble)
```

If we explore the following list, we can see that most 

```{r}
str(gh_repos)
data("gh_repos") #run in to the environment
gh_repos[[1]] #now R environment shows something

#give a look at list
View(gh_users)
```

extract information from lists

### warmup challenge: extract information full name of first repo
```{r}
gh_repos %>% 
  map_chr(list(1,"full_name"))

gh_repos %>% map_chr(c(1,3))
```

We can use the position to find the second element in the second level
```{r}
gh_repos %>%
  map_chr(c(2,2)) 
```

But, we can also specify the name of the variable
```{r}
gh_repos %>% 
  map_chr(list(2,"name"))
```

The notation in `[ ]` for each element, should be something like the following, but map functions allows you to do the operation for all elements at once
```{r}
gh_repos[[2]][[2]][["name"]]
```

### challenge: get the owner id of fourth repo for each user (its of the type integer)
```{r}
gh_repos %>% 
  map_int(list(4,"owner","id"))
```

#### dataframes

how to look for the first level of a list for `str`
```{r}
str(gh_repos, max.level = 1)
```

Now, we have to create names to be able to create a dataframe
```{r}
u_id <- gh_repos %>% 
  map_int(list(1, "owner", "id"))

gh_repos_named <- gh_repos %>% 
  set_names(u_id)  # we need to set names to these lists

str(gh_repos_named, max.level = 1)
```

To create a dataframe we have to give names to the columns in that tibble
```{r}
gh_repos_df <- gh_repos_named %>% 
  enframe("user_id", "repository_stuff")

gh_repos_df
```

### challenge: now lets try to create an enframe dataframe but identifying by their owner login 

```{r}
login <- gh_repos %>% map_chr(list(1,"owner", "login"))
gh_repos_login <- gh_repos %>% set_names(login)
gh_login_df <- gh_repos_login %>% enframe("user_login", "repo_info")
gh_login_df
```

From this we can use other functions, such as measure the length, or other functions

```{r}
gh_login_df %>% 
  filter(user_login %in% c("masalmon", "leeper")) %>% 
  mutate(number_of_repositories = length(repo_info))
# here, length is unable to quantify what is inside the list

gh_login_df %>% 
  filter(user_login %in% c("masalmon", "leeper")) %>% 
  mutate(number_of_repositories = map_int(repo_info,
         length))
```

Now, from repurrrsive lets use one of the lists `got_chars`
```{r}
map(got_chars,"aliases")
map(got_chars,"name")
```

Lets make a list of these two things together
```{r}
aliases_name <- set_names(map(got_chars, "aliases"),
                          map(got_chars, "name"))

# see the difference between these two
length(aliases_name)
map_int(aliases_name, length)
```

create a function
```{r}
collapse_by_comma <- function(vec_of_strings){
  paste(vec_of_strings, collapse = ", ")
}
collapse_by_comma(c("first string", "second string"))

### different ways to apply the same function, take the first element of a list, and apply a function to that 

map(aliases_name, collapse_by_comma)
map(aliases_name, 
    function(v_o_s) paste(v_o_s, collapse = ", "))
map(aliases_name, paste, collapse = ", ")
map(aliases_name, ~ paste(.x), collapse = ", ")
```

