---
title: "cm107_my_notes"
author: "Santiago David"
date: '2017-11-14'
output: html_document
---

To practice with Shiny apps, we need several packages
```{r}
#install.packages("shiny")
library(shiny)
```

We can start by exploring example
```{r}
runExample("01_hello")
```

We can stop the file from executing... this is a browser working in your computer

Now, we need to prepare a user interface and a server...

```{r}
my_ui <- fluidPage()
my_server <- function(input, output){
  
}
shinyApp(ui = my_ui, server = my_server)
```

We have to save this file in a separate R file in a specific folder, and we name the file `App.R`. When we do that, there will be a `Run` option in the file.
This websites are hoted locally, there are still no in the internet.
This files understand html, so we can use html notation

The inputs come from the user interface, and become available to the server, by using an argument `input`. We have to use `render` to produce some output from the input and save it. We always need to close the render with a curly parenthesis {}. So basically what this is doing, is taking inputs sending them to the server and asking it to do something.
So, Shiny is basically creating that interface for html language based on the information we are sending...

Shiny focuses on the data analysis and provide a structure for displaying that rely on html

Its recommended to separate the ui.R file from the server.R file

#### cm108 - 16 Novemeber 2017

The UI displays the outputs
The server takes inputs and create some outputs

