---
title: "cm103"
author: "Santiago David"
date: '2017-10-31'
output: github_document
---

### Indexing a vector, purrr, map_functions

An atomic vector is the one in which all elements are of the same type (not a list of mixed elements), for example

```{r}
# trick if you sorround the vector on parenthesis, it prints the content
(v_log <- c(TRUE, FALSE, FALSE, TRUE))
(v_int <- 1:4)
(v_double <- 1:4 * 1.2)
(v_char <- letters[1:4])
typeof(v_double)
typeof(v_char)
```

Now, we can use a sort of filter function to test a statement on the data, 
or index the data by logical statement
```{r}
v_char[c(FALSE, FALSE, TRUE, TRUE)]
v_char[v_log]
```
How do we index by position?
```{r}
v_double[2:3]
v_double[1]
v_char[-4]
```

We can mess up a vector by coercing it to a different type
```{r}
v_int[1]
typeof(v_int[1])
v_doub_copy <- v_double
str(v_doub_copy)
#now if we modified one value to character is gonna change the whole type of the vector
v_doub_copy[3] <- "uhoh"
str(v_doub_copy)
```

lists
```{r}
(x <- list(1:3, c("four", "five")))
(y <- list(logical = TRUE, integer = 4L, double = 4*1.2, character = "character"))
(z <- list(letters[26:22], trascendental = c(pi, exp(1)), f = function(x)x^2))
```

#### challenge, in list y, pull out the second and third elements, in list z pull out all the trascendental numbers

```{r}
y[2:3]
z[2]
z["trascendental"]
z$trascendental
```

Now, if we look at the typeof we see that we are only looking at the type of the variable, but if we want to access the content, we need [[...]]
```{r}
typeof(z["trascendental"])
z[["trascendental"]]
y[["double"]]
y["double"]
```

#### for loops

Consider how to square the integers 1 through n, with n <- 5
```{r}
n <- 5
res <- rep(NA_integer_, n)
for (i in seq_len(n)) {
res[i] <- i ^ 2  
}
res
```

another option is just
```{r}
(1:n)^2
```

now, we take another vector
```{r}
exp(v_double)
(l_doub <- as.list(v_double))
```
now, what if we want to elevated all the elements in l_doub
```{r}
exp(l_doub) #not possible
```

Let's start with `purrr`
```{r}
library(tidyverse)
library(purrr)
library(stringr)
```

we can use the function `map_dbl`
```{r}
map_dbl(l_doub, exp)
```

### challenge, use a `map_` and `stringr` function to split all the fruits that have two words in them to two separate character strings
```{r}
fruit
str(fruit)

fruit %>% 
str_subset(" ") #not finished
```

Class solution
```{r}
#map_chr(fruit, str_split, " ")
fruit %>% 
  map(str_split, " ")

fruit %>% 
  map_chr(str_to_upper)
```

## Simplifying data from a [list](https://htmlpreview.github.io/?https://github.com/JoeyBernhardt/stat545-purrr/blob/master/purrr-lesson.html)
```{r}
library(repurrrsive)
library(listviewer)
library(jsonlite)
library(dplyr)
library(tibble)
library(purrr)
```

We will use `gh_users` as an example to extract and index data from a list
```{r}
str(gh_users)
jsonedit(gh_users) #allows to see the info in a viewer
```

start using map_ functions
```{r}
map(gh_users, "login") #by name
gh_users %>% 
  map(18) # by position
```

We want to pull out multiple things
```{r}
x <- map(gh_users, `[`, c("login", "name", "id", "location"))
x #now we can index only the elements that we want
```

```{r}
y <- map(gh_users, magrittr::extract, c("login", "name"))
y
```

Try more map functions before next class... for example for calculating values such as `mean`

