---
title: "cm101"
author: "Santiago David"
date: '2017-10-24'
output: github_document
---

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(singer))
suppressPackageStartupMessages(library(stringr))
```

# R for strings

```{r}
my_string <- "Surrounded by quotes, any text"
my_string

my_string2 <- 'if you use single quotes, "you have to close with single quotes". '
my_string2
```

A slash \ is a special character that can't be included by itself, because it has a meaning in r language

```{r}
my_string3 <- " A slash \ "
my_string3
my_string3 <- " A slash \\"
```

### Challenge: How do I "string" a single slash?

```{r}
my_string4 <- `my_string`
my_string5 <- 'my_string'
my_string6 <- "my_string"
my_string4 == my_string5
my_string5 == my_string6
```

this is called vec quotes ` `, they can surround strange and conflictive characters

```{r}
places <- unique(singer_locations$city) #get all cities
str(places)
```

nchar() takes a vector of strings and return its length

```{r}
nchar(c("city", "double", "a very long string"))
nchar(c("", NA, " ", "    ", " a "))
nchar(c("  ", "\t")) # a tab which is a special character, and \t which simbolizes a tab
```

## The tools of the trade

```{r}
devtools::install_github("tidyverse/glue", force = TRUE)
library(glue)
```

glueing `glue()`

In base R we use paste(), but we can also use glue(), this function is executed in the global environment and returns the value

```{r}
My_name <- "Santiago"
My_Age <- 31
My_color <- "blue"

glue("My name is {My_name}", "My age two years ago was {My_Age - 2}")
```
```{r}
paste0("My name is ",My_name,", My age two years ago was ",My_Age - 2)
```

Indentation
```{r}
glue("
     My name is
          {My_name},
My age two years ago was
    {My_Age - 2}")
```

### Challenge

Phrase: Today is Wednesday. October 29. 1986 and the sun shines
Sys.Date()
glue()

```{r}
day <- Sys.Date()
glue("Today is {day}. and the sun shines")
```

We can make it faster by creating an object, class solution!
```{r}
phrase <- "Today is {Today_date}. and the sun shines"
glue(phrase,
     Today_date = Sys.Date())
```

change format
```{r}
glue(phrase,
     Today_date = format(Sys.Date(), "%A, %B, %d, %Y"))
```

`glue_data` understand data.frames, so we can extract data for all values in a dataframe and "glue" them to a string

```{r}
singer_locations %>% 
  glue_data("{artist_name} is from {city}.") %>% 
  head()
```

### stringr

Let's check at the possible options
```{r}
getNamespaceExports("stringr")
```

str_view provides all strings that matches a certain pattern

```{r}
getNamespaceExports("stringr") %>% 
  str_view("str_")
```

```{r}
getNamespaceExports("stringr") %>% 
  str_subset("str_")
```

### Challenge

Let's get all the places containing string "york"

```{r}
places %>% 
  str_subset("york")
```

But, what if there are more cases with "York", for that we can use `str_to_lower()` which is looking for capital and lower case

```{r}
places %>% 
  str_to_lower() %>% 
  str_subset("york")
```

Bue we lose the ability to see the original string

## REgular Expressions 
sometimes, we want to find things that are similar to a certain pattern that we define

things inside [ ] define any of the two options, 
```{r}
places %>% 
  str_subset("[Yy]ork")
```

another funtion is coll(), which looks for more options, as for example not considering the case
```{r}
places %>% 
  str_subset(coll("york", ignore_case = TRUE))
```

so far, we also getting "yorkshire", so another option to only get new york is using `$` which define the end of a string or `^` which define the start
```{r}
places %>% 
  str_subset("York$")

places %>% 
  str_subset("^York")
```

another option `:digit:` will define any digits in a string
Challenge: lets get all places containing digits

```{r}
"[:digit:]"
places %>% 
  str_subset("[:digit:]")
```

this is good to start cleaning a dataset, since we can identify mistakes and info that is not a city

The point `.` is a wildcard that allows to identify any value with that condition
```{r}
places %>% 
  str_subset("B.a")
places %>% 
  str_subset("B.a?")
places %>% 
  str_subset("B.a+")
places %>% 
  str_subset("B.a*") #* means zero one or more
places %>% 
  str_subset("[Nn]ew [Yy]ork")
```

There are a lot more regular expressions `str_`

## HW use regular expressions and functions to clean up city places in dataframe
