---
title: "Scraping Work_gdvr example"
author: "Santiago David"
date: '2017-11-30'
output: github_document
---

Packages
```{r}
library(tidyverse)
library(magrittr)
library(purrr)
library(glue)
library(stringr)
library(rvest)
library(xml2)
```

Titles

```{r}
url_titles <- "https://www.musixmatch.com/artist/Straight-Line-Stitch#"

page_title <- read_html(url_titles)
page_title
```

The CSS selector
```{r}
page_title %>%
  html_nodes(".title")
```

```{r}
page_title %>%
  html_nodes(".title") %>%
  html_text()
```

We want the lyrics
```{r}
SLS_df <- data_frame(Band = "Straight Line Stitch",
                     Title = page_title %>%
                       html_nodes(".title") %>%
                       html_text())
```

```{r}
SLS_lyrics <- SLS_df %>% mutate(Link = glue('https://www.musixmatch.com/lyrics/{Band}/{Title}') %>%
                           str_replace_all(" ","-"))
```

```{r}
page_title %>%
  html_nodes(".title") %>%
  html_attrs() %>%
  glimpse()
```

```{r}
page_title %>%
  html_nodes(".title") %>%
  html_attrs() %>%
  map_chr("href")
```

using rvest
```{r}
page_title %>%
  html_nodes(".title") %>%
  html_attr("href")
```

```{r}
SLS_df %<>%
  mutate(Link = page_title %>%
  html_nodes(".title") %>%
  html_attr("href"))
```

And purrr
```{r}
url_song <- glue("https://www.musixmatch.com{SLS_df$Link[1]}")

url_song
```

```{r}
url_song %>%
  read_html() %>%
  html_nodes(".mxm-lyrics__content") %>%
  html_text()
```

collapse() and glueing() the part that is missing in the link, which is the www...

```{r}
get_lyrics <- function(link){
  glue("https://www.musixmatch.com{link}") %>%
   read_html() %>%
   html_nodes(".mxm-lyrics__content") %>%
   html_text() %>%
   collapse(sep = "\n") %>%
    return()
}
```

```{r}
SLS_df$Link[3] %>%
  get_lyrics()
```

```{r}
SLS_df %<>%
  mutate(Lyrics = map_chr(Link, get_lyrics))
```

Another artist
```{r}
AH_url <- "https://www.musixmatch.com/artist/Angel-Haze"

AH_lyrics <- data_frame(Band = "Angel Haze",
                        
                         Title = AH_url %>%
                          read_html() %>%
                           html_nodes(css = ".title") %>%
                           html_text(),
                        
                         Link = AH_url %>%
                          read_html() %>%
                           html_nodes(css = ".title") %>%
                          html_attr("href")
                        )
```


Getting the lyrics, we use a map_chr() function to run the function for all the titles
```{r}
AH_lyrics %>% 
  mutate(Lyrics = map_chr(Link, get_lyrics))
```

Another example with a more complicated scenario
Another artist called A tribe called red, in which some lyrics doesn't exist

```{r}
ATCR_url <- "https://www.musixmatch.com/artist/A-Tribe-Called-Red"

ATCR_lyrics <- data_frame(Band = "A Tribe Called Red",
                        
                         Title = ATCR_url %>%
                          read_html() %>%
                           html_nodes(css = ".title") %>%
                           html_text(),
                        
                         Link = ATCR_url %>%
                          read_html() %>%
                           html_nodes(css = ".title") %>%
                          html_attr("href")
                        )
```

Now, the previous function to get lyrics will not work, because there are missing lyrics
```{r}
ATCR_lyrics %>% 
  mutate(Lyrics = map_chr(Link, get_lyrics))
```
We need to create a new function that uses `purrr::possibly` to assign an NA when another function doesn't work

```{r}
get_lyrics_safe <- purrr::possibly(get_lyrics, "None available")
```

but instead of using `map_chr` we just use `map`
```{r}
ATCR_lyrics %>% 
  mutate(Lyrics = map(Link, get_lyrics_safe)) %>% 
  unnest()
```


