---
title: "cm12"
author: "Santiago David"
date: '2017-10-12'
output: github_document
---

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(gapminder))
```

More on typeof/class

```{r}
typeof("A") #character
typeof(1) #numeric
typeof(gapminder)
typeof(mtcars) #
typeof(lm) # function
typeof(TRUE) #logical (Boolean)
typeof(c) #builtin

class("A")
class(1)
class(gapminder)
class(mtcars)
class(lm)
class(c)
```
We can also create a list
```{r}
x <- list(number_of_cats = 1, cat_color = "red")
typeof(x)
class(x)
```

Other type is matrix, matrices need the same type of objects, if there are characters, the matrix will force everything to be a character

```{r}
my_mat <- matrix(c(1,2,3,4), nrow = 2, byrow = TRUE)
my_mat
```

looking at singer 
```{r}
library(singer)
data("singer_locations")
```

# change a type of variable in singer_locations from its type
```{r}
copy <- singer_locations
str(singer_locations)
as.numeric(copy$track_id[1])

#also
glimpse(copy)
#now we change that type
copy[1,8] <- "long"
```

### Factors

```{r}
glimpse(gapminder)
```

Factor is a tabular way of looking at strings??, there is a relationship between a number and a level, so what it means for that number

Library to inspect factors
```{r}
#install.packages("forcats")
library(forcats)
library(singer)
```

lets work on singer database

```{r}
data("singer_locations")
glimpse(singer_locations)
```

Transforming into a factor

```{r}
sl <- singer_locations %>% 
  mutate(artist_factor =  as_factor(artist_name))
glimpse(sl)
```

we can also do it using base R
```{r}
sl <- sl %>% 
  mutate(artist_factor_base = factor(artist_name))
glimpse(sl)
```

```{r}
x <- c("a", "b", "c")
as_factor(x)
factor(x)
```

When we use base R to asign factor levels, R will do it alphabetically, which might not be what we need, `as_factor` on the other hand maintains the order we have.

Now, we can start playing with the factors

```{r}
sl %>% 
  mutate(top_an = fct_lump(artist_factor, n = 10)) %>%  #lumping function to have one group of other
  count(top_an) %>%
  arrange(-n)
```


### Homework: plot mean duration by artist with and without lumping

First, we can try without...

```{r}
sl %>% 
  group_by(artist_name) %>% 
  summarise(mean_duration = mean(duration)) %>% 
  ggplot(aes(x = artist_name, y = mean_duration))
```

Now, we can use the factor to create the plot

```{r}
sl %>% 
  mutate(top_an = fct_lump(artist_factor, n = 5)) %>% 
  group_by(top_an) %>% 
  summarise(mean_duration = mean(duration)) %>% 
  ggplot(aes(x = top_an, y = mean_duration)) +
  geom_point()
```



### now we can use cities
```{r}
sl <- sl %>% 
  mutate(city_factor = as.factor(city))
glimpse(sl)
```

In case there are NA's conflicting the levels, we can use `ifelse` to name those missing info

```{r}
sl <- sl %>% 
  mutate(city_factor = ifelse(is.na(city),
                              "Missing information", 
                              city),
  city_factor = as_factor(city_factor))

```

Now, we look 
```{r}
sl %>% 
  mutate(top_cities = fct_lump(city_factor, n = 7)) %>% 
  count(top_cities)
```

### homework challenge
use `forcats` to clean up the city stuff. So, basically rename factors
forcats::as_factors

```{r}
sl %>% 
  mutate(city_factor = fct_recode(city_factor,
                                  "New York, NY" = "NY - New York City",
                                  "Los Angeles, CA" = "California - LA")) %>% 
  mutate(top_cities = fct_lump(city_factor, n = 7)) %>% 
  count(top_cities)
```



### Reading and writing to file

`read_csv` which is different than `read.csv`, and apparently make the process easier by identifying the variables better

`readxl` is another useful function 

```{r}
#install.packages("readxl")
```

`readxl::read_xlsx()` directly takes that file form excel

`write_csv` do the opposite, write the file

